Master Thesis

### 2013-06-04 my-work-rewritten.tex
Prettier implementation of lexicons

(def string-sig
  (ors (nonlex-sigr {'++ (-> 'Str 'Str 'Str)})
       (unitypedr 'Str)))

@@

Our representation of lexical constants was purposefully engineered so
as to make the lexical entry directly accessible to the lexicon
implementation. This way, the lexicon author can inspect the hypertag
and the wordform of the lexical entry, the specifier of the concrete
constant that it is mapping right now and even the type she assigned to
the constant in the abstract signature to allow her to choose the
correct object term. Furthermore, since lexicons are still
relations/sets, we can compose them together by disjunction. See
Figure~\ref{lex-impl} for the implementation of the lexicalized versions
of the $\mathcal{L}_{Syntax}$ and $\mathcal{L}_{Sem}$ lexicons.

\begin{figure}
  \centering
\begin{verbatim}
(defn syntax-lexo
  [synt-constant string-term]
  (with-sig-consts string-sig
    (l/fresh [hypertag string-constant]
      (has-hypertago synt-constant hypertag)
      (share-lex-entryo synt-constant string-constant)
      (string-sig string-constant)
      (fs-assigne hypertag string-term
                  {:head {:cat "n"}}
                  ,(rt string-constant)
                  {:head {:cat "v"
                          :trans "true"}}
                  ,(rt (ll [x y] (++ (++ x string-constant) y)))
                  {:head {:cat "det"}}
                  ,(ll [x R] (R (++ ua-stx-constant x)))))))

(defn sem-lexo
  [synt-constant sem-term]
  (with-sig-consts sem-sig
    (l/fresh [hypertag sem-constant]
      (has-hypertago synt-constant hypertag)
      (l/conde [(share-lex-entryo synt-constant sem-constant)
                (sem-sig sem-constant)
                (fs-assigne hypertag sem-term
                            {:head {:cat "n"}}
                            ,(rt sem-constant)
                            {:head {:cat "v"
                                    :trans "true"}}
                            ,(rt sem-constant))]
               [(fs-assigne hypertag sem-term
                            {:head {:cat "det"
                                    :det_type "indef"}}
                            ,(rt (ll [p q] (exists? (il [x] (and? (p x)
                                                                  (q x))))))
                            {:head {:cat "det"
                                    :det_type "def"}}
                            ,(rt (ll [p q] (forall? (il [x] (imp? (p x)
                                                                  (q x)))))))]))))

\end{verbatim}
  \caption{\label{fig:lex-impl} The implementation of the lexicons from
    subsection~\ref{ssec:acg}. NOTE: This is too verbose both to put on
    the paper and also to code. These examples demonstrates two common
    and related shapes of lexicons. The $\mathcal{L}$(\_{Syntax}) and
    the first half of the $\mathcal{L}$(\_{Sem}) are lexicalized, in the
    sense that the object term contains a lexical constant with the same
    lexical entry as the abstract constant.\footnote{This notion of
      ``lexicalized'' is quite different to the one I use in the rest of
      the thesis so far (which roughly means ``generated by a lexical
      database'') and also different from the one used by Kanazawa
      (FIX!).} This means that before constructing the object term, we
    have to acquire the object constant that is contained in the term
    and then we just fit it in of the recipes, which we choose according
    to the hypertag. The second half of $\mathcal{L}$(\_{Sem}) simply
    assigns object terms based on the hypertag. A more concise and
    desirable implementation would allow us simply write the pairs of
    hypertags and object terms (recipes encoded as lambdas). My plan is
    to have exactly this, but I have not thought yet about how to split
    the various aspects into different reusable abstractions.}
\end{figure}

>>>>

(def string-sig
  (ors (nonlex-sigr {'++ (-> 'Str 'Str 'Str)
                     'empty-str 'Str})
       (unitypedr 'Str)))

@@

Our representation of lexical constants was purposefully engineered so
as to make the lexical entry directly accessible to the lexicon
implementation. This way, the lexicon author can inspect the hypertag
and the wordform of the lexical entry, the specifier of the concrete
constant that it is mapping right now and even the type she assigned to
the constant in the abstract signature to allow her to choose the
correct object term. Furthermore, since lexicons are still
relations/sets, we can compose them together by disjunction. See
Figure~\ref{fig:lex-impl} for the implementation of the lexicalized
versions of the $\mathcal{L}_{Syntax}$ and $\mathcal{L}_{Sem}$ lexicons.

\begin{figure}
  \centering
\begin{verbatim}
(def syntax-lexo
  (with-sig-consts string-sig
    (lexicalizer string-sig
                 (ht-lexiconr {{:head {:cat "n"}}
                               ,(rt (ll [_] _))
                               {:head {:cat "v"
                                       :trans "true"}}
                               ,(rt (ll [_ x y] (++ (++ x _) y)))
                               {:head {:cat "det"}}
                               ,(rt (ll [_ x R] (R (++ _ x))))}))))

(def sem-lexo
  (with-sig-consts sem-sig
    (orr (lexicalizer sem-sig
                      (ht-lexiconr {{:head {:cat "n"}}
                                    ,(rt (ll [_ x] (_ x)))
                                    {:head {:cat "v"
                                            :trans "true"}}
                                    ,(rt (ll [_ s o] (_ s o)))}))
         (ht-lexiconr {{:head {:cat "det"
                               :lemma "un"}}
                       ,(rt (ll [P Q] (exists? (il [x] (and? (P x)
                                                             (Q x))))))
                       {:head {:cat "det"
                               :lemma "chaque"}}
                       ,(rt (ll [P Q] (forall? (il [x] (imp? (P x)
                                                             (Q x))))))}))))
\end{verbatim}
  \caption{\label{fig:lex-impl} The implementation of the
    $\mathcal{L}$(\_{Syntax}) and $\mathcal{L}$(\_{Sem}) lexicons from
    subsection~\ref{ssec:acg}.}
\end{figure}

 NOTE: This is too verbose both to put on the paper and also to
 code. These examples demonstrate two common and related shapes of
 lexicons. The $\mathcal{L}$(\_{Syntax}) and the first half of the
 $\mathcal{L}$(\_{Sem}) are lexicalized, in the sense that the object
 term contains a lexical constant with the same lexical entry as the
 abstract constant.\footnote{This notion of ``lexicalized'' is quite
   different to the one I use in the rest of the thesis so far (which
   roughly means ``generated by a lexical database'') and also different
   from the one used by Kanazawa (FIX!).} This means that before
 constructing the object term, we have to acquire the object constant
 that is contained in the term and then we just fit it in of the
 recipes, which we choose according to the hypertag. The second half of
 $\mathcal{L}$(\_{Sem}) simply assigns object terms based on the
 hypertag. A more concise and desirable implementation would allow us
 simply write the pairs of hypertags and object terms (recipes encoded
 as lambdas). My plan is to have exactly this, but I have not thought
 yet about how to split the various aspects into different reusable
 abstractions.
